---
title: "Forecast - JyskeBank"
author: "Andreas Borup Joergensen, Mette Koch Moeller, Robert Hoestrup"
Date: "04-03-2021"
output:
  html_notebook:
    code_folding: show
    df_print: paged
    theme: flatly
    toc: yes
    toc_depth: 4
    toc_float:
      collapsed: no
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '2'
---

```{r setup, include=FALSE}
### Generic preamble
rm(list=ls())
Sys.setenv(LANG = "en")
options(scipen = 5)

library(knitr)
knitr::opts_chunk$set(warning=FALSE,
                     message=FALSE,
                     comment=FALSE, 
                     fig.align="center"
                     )
```

Downloading the packages
```{r}
library(tidyverse) #Collection of all the good stuff like dplyr, ggplot2 ect.
library(magrittr) #For advanced piping
library(keras) #For ANN models
library(kerastuneR) #For hyper tuning ANN models
library(tidymodels)
library(ggpubr) #package that helps mixing multiple plots on the same page
```

```{r}
data <- read_csv("https://raw.githubusercontent.com/andreasbj77/Master-thesis/main/Data/Final_data/DataCreationJyskeBank.csv")
```

The data contains the date, the response variable and the features we have decided to use.
```{r}
data %>% glimpse()
```

# Pre-processing
_____
The variable is scaled and made numeric. This is done because its generally easier for RNNs to learn from scaled data. The scaling attribute attr is saved in order to be able to unscale the data later.

```{r}
data$JyskeBank %<>% scale(center = FALSE, scale = TRUE)
data$SP500 %<>% scale(center = FALSE, scale = TRUE)

att_JyskeBank <- attr(data$JyskeBank, 'scaled:scale')

data$JyskeBank %<>% as.numeric()
data$SP500 %<>% as.numeric()
```

We check the data to make sure it is all scaled properly
```{r}
data %>% glimpse()
```

The variables needed to forecast JyskeBank is selected from the complete data set
```{r}
data_JyskeBank <- data
```

## Out-of-Sample-Testing

The data is split into a training, testing, and forecasting set. The periods chosen are the 2010-2017 for training, 2018 for testing and 2019 for forecasting.

```{r}
forecasting <- data_JyskeBank %>% filter(Date >= "2018-12-29")
```


The data sets are divided into the response and explanatory variables. Here a lead of JyskeBank acts as the response variable.

```{r}

#Forecast data
x_forecast <- forecasting %>% 
  select(-Date) %>% 
  as.matrix()

y_forecast <- forecasting %>%
  select(-Date) %>%
  mutate(JyskeBank = JyskeBank %>% lead(1)) %>%
  mutate(JyskeBank = ifelse(is.na(JyskeBank), lag(JyskeBank, 1), JyskeBank)) %>%
  select(JyskeBank) %>% 
  as.matrix()
```

```{r}
n_input = 5  # amount of timesteps in the model
n_features = ncol(data_JyskeBank) - 1 # amount of explanatory variables
b_size = 64 # amount of data points going through at the same time
step_forecast = (nrow(forecasting)/b_size) %>% round(0) # Steps is equals to the amount that sends all data through the model once per epoch

# Generator for forecasting
forecast_generator = timeseries_generator(x_forecast,
                                 y_forecast,
                                 length = n_input,
                                 batch_size = b_size)
```


# Loading models

```{r}
model_FNN <- load_model_hdf5("E:/Hyper_test/JyskeBank/model_ANN_JyskeBank")
model_RNN <- load_model_hdf5("E:/Hyper_test/JyskeBank/model_RNN_JyskeBank")
model_LSTM <- load_model_hdf5("E:/Hyper_test/JyskeBank/model_LSTM_JyskeBank")
model_biLSTM <- load_model_hdf5("E:/Hyper_test/JyskeBank/model_biLSTM_JyskeBank")
model_GRU <- load_model_hdf5("E:/Hyper_test/JyskeBank/model_GRU_JyskeBank")
model_biGRU <- load_model_hdf5("E:/Hyper_test/JyskeBank/model_biGRU_JyskeBank")
```



# Results

```{r warning=FALSE}
FNN_prediction <- model_FNN %>% predict(x_forecast)
RNN_prediction <- model_RNN %>% predict_generator(forecast_generator, steps = step_forecast)
LSTM_prediction <-model_LSTM %>% predict_generator(forecast_generator, steps = step_forecast)
biLSTM_prediction <-model_biLSTM %>% predict_generator(forecast_generator, steps = step_forecast)
GRU_prediction <-model_GRU %>% predict_generator(forecast_generator, steps = step_forecast)
biGRU_prediction <-model_biGRU %>% predict_generator(forecast_generator, steps = step_forecast)
```

```{r}
FNN_prediction_unscale <- FNN_prediction * att_JyskeBank
RNN_prediction_unscale <- RNN_prediction * att_JyskeBank
LSTM_prediction_unscale <- LSTM_prediction * att_JyskeBank
biLSTM_prediction_unscale <- biLSTM_prediction * att_JyskeBank
GRU_prediction_unscale <- GRU_prediction * att_JyskeBank
biGRU_prediction_unscale <- biGRU_prediction * att_JyskeBank

y_forecast_unscale <- y_forecast * att_JyskeBank
```


```{r}
start <- n_input + 1

eval_FNN <- tibble(
  Date = forecasting %>% pull(Date),
  truth = y_forecast_unscale,
  prediction = FNN_prediction_unscale
)

eval_RNN <- tibble(
  Date = forecasting[start:nrow(forecasting),] %>% pull(Date),
  truth = y_forecast_unscale[start:nrow(forecasting),],
  prediction = RNN_prediction_unscale
)

eval_LSTM <- tibble(
  Date = forecasting[start:nrow(forecasting),] %>% pull(Date),
  truth = y_forecast_unscale[start:nrow(forecasting),],
  prediction = LSTM_prediction_unscale
)

eval_biLSTM <- tibble(
  Date = forecasting[start:nrow(forecasting),] %>% pull(Date),
  truth = y_forecast_unscale[start:nrow(forecasting),],
  prediction = biLSTM_prediction_unscale
)

eval_GRU <- tibble(
  Date = forecasting[start:nrow(forecasting),] %>% pull(Date),
  truth = y_forecast_unscale[start:nrow(forecasting),],
  prediction = GRU_prediction_unscale
)

eval_biGRU <- tibble(
  Date = forecasting[start:nrow(forecasting),] %>% pull(Date),
  truth = y_forecast_unscale[start:nrow(forecasting),],
  prediction = biGRU_prediction_unscale
)
```

```{r}
RMSE_FNN <- eval_FNN %>% rmse(as.numeric(truth), as.numeric(prediction))
MAPE_FNN <- eval_FNN %>% mape(as.numeric(truth), as.numeric(prediction))
RMSE_FNN %<>% rename(RMSE = .estimate)
MAPE_FNN %<>% rename(MAPE = .estimate)
FNN <- cbind(RMSE_FNN, MAPE_FNN)
FNN %<>% select(RMSE, MAPE) %>% mutate(Model = "FNN")

RMSE_RNN <- eval_RNN %>% rmse(as.numeric(truth), as.numeric(prediction))
MAPE_RNN <- eval_RNN %>% mape(as.numeric(truth), as.numeric(prediction))
RMSE_RNN %<>% rename(RMSE = .estimate)
MAPE_RNN %<>% rename(MAPE = .estimate)
RNN <- cbind(RMSE_RNN, MAPE_RNN)
RNN %<>% select(RMSE, MAPE) %>% mutate(Model = "RNN")

RMSE_LSTM <- eval_LSTM %>% rmse(as.numeric(truth), as.numeric(prediction))
MAPE_LSTM <- eval_LSTM %>% mape(as.numeric(truth), as.numeric(prediction))
RMSE_LSTM %<>% rename(RMSE = .estimate)
MAPE_LSTM %<>% rename(MAPE = .estimate)
LSTM <- cbind(RMSE_LSTM, MAPE_LSTM)
LSTM %<>% select(RMSE, MAPE) %>% mutate(Model = "LSTM")

RMSE_biLSTM <- eval_biLSTM %>% rmse(as.numeric(truth), as.numeric(prediction))
MAPE_biLSTM <- eval_biLSTM %>% mape(as.numeric(truth), as.numeric(prediction))
RMSE_biLSTM %<>% rename(RMSE = .estimate)
MAPE_biLSTM %<>% rename(MAPE = .estimate)
biLSTM <- cbind(RMSE_biLSTM, MAPE_biLSTM)
biLSTM %<>% select(RMSE, MAPE) %>% mutate(Model = "biLSTM")

RMSE_gru <- eval_GRU %>% rmse(as.numeric(truth), as.numeric(prediction))
MAPE_gru <- eval_GRU %>% mape(as.numeric(truth), as.numeric(prediction))
RMSE_gru %<>% rename(RMSE = .estimate)
MAPE_gru %<>% rename(MAPE = .estimate)
GRU <- cbind(RMSE_gru, MAPE_gru)
GRU %<>% select(RMSE, MAPE) %>% mutate(Model = "GRU")

RMSE_bigru <- eval_biGRU %>% rmse(as.numeric(truth), as.numeric(prediction))
MAPE_bigru <- eval_biGRU %>% mape(as.numeric(truth), as.numeric(prediction))
RMSE_bigru %<>% rename(RMSE = .estimate)
MAPE_bigru %<>% rename(MAPE = .estimate)
biGRU <- cbind(RMSE_bigru, MAPE_bigru)
biGRU %<>% select(RMSE, MAPE) %>% mutate(Model = "biGRU")
```

The resulting MAPE and RMSE for each of the models are combined in a single data frame
```{r}
Results <- rbind(FNN, RNN, LSTM, biLSTM, GRU, biGRU)
```

## Testing results

We arrange the results at ascending RMSE to see which model performs best
```{r}
Results %>% select(Model, RMSE, MAPE) %>% arrange(RMSE)
```

We also plot all the models' forecast to see how they perform over time
```{r fig.height=20, fig.width=15}
plot_FNN <- eval_FNN %>% 
  pivot_longer(-Date) %>%
  ggplot(aes(x = Date, y = value, col = name)) +
  geom_line() +
  ggtitle("FNN")

plot_RNN <- eval_RNN %>% 
  pivot_longer(-Date) %>%
  ggplot(aes(x = Date, y = value, col = name)) +
  geom_line() +
  ggtitle("RNN")

plot_LSTM <- eval_LSTM %>% 
  pivot_longer(-Date) %>%
  ggplot(aes(x = Date, y = value, col = name)) +
  geom_line() +
  ggtitle("LSTM")

plot_biLSTM <- eval_biLSTM %>% 
  pivot_longer(-Date) %>%
  ggplot(aes(x = Date, y = value, col = name)) +
  geom_line() +
  ggtitle("biLSTM")

plot_GRU <- eval_GRU %>% 
  pivot_longer(-Date) %>%
  ggplot(aes(x = Date, y = value, col = name)) +
  geom_line() +
  ggtitle("GRU")

plot_biGRU <- eval_biGRU %>% 
  pivot_longer(-Date) %>%
  ggplot(aes(x = Date, y = value, col = name)) +
  geom_line() +
  ggtitle("biGRU")

plot <- ggarrange(plot_FNN, plot_RNN, plot_LSTM, plot_biLSTM, plot_GRU, plot_biGRU, ncol = 1)
annotate_figure(plot, top = text_grob("Accuracy of forecast", size = 15))
```

```{r}
Results %>% select(Model, RMSE, MAPE) %>% write.csv("forecast_JyskeBank")
```

