---
title: "Econometrics - Genmab"
author: "Andreas Borup Joergensen, Mette Koch Moeller, Robert Hoestrup"
date: "04-03-2021"
output:
  html_notebook:
    code_folding: show
    df_print: paged
    theme: flatly
    toc: yes
    toc_depth: 4
    toc_float:
      collapsed: no
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '2'
---
```{r setup, include=FALSE}
### Generic preamble
rm(list=ls())
Sys.setenv(LANG = "en")
options(scipen = 5)

library(knitr)
knitr::opts_chunk$set(warning=FALSE,
                     message=FALSE,
                     comment=FALSE, 
                     fig.align="center"
                     )
```

Downloading the packages
```{r}
library(tidyverse) #Collection of all the good stuff like dplyr, ggplot2 ect.
library(magrittr) #For advanced piping
library(ggpubr) #package that helps mixing multiple plots on the same page
library(wooldridge)
library(car)
library(lmtest)
library(sandwich)
library(quantmod)
library(tseries)
library(urca)
library(vars)
library(forecast)
library(nortsTest)
library(yardstick)
library(midasr) # Brugt til static forecasting
#'forecast', 'midasr', 'nortsTest', 'quantmod', 'sandwich', 'tseries', 'urca', 'vars', 'wooldridge'
```

Loading data from GitHub

```{r} 
data <- read.csv("https://raw.githubusercontent.com/andreasbj77/Master-thesis/main/Data/Final_data/DataCreationGenmab.csv")
```

# Preprocessing
We split up our data set into training, testing, training+testing and forecasting. Splitting up the data makes it possible for us to make out-of-sample forecasting. The training+testing data frame is created so we can make one step ahead forecasts in the testing period.
```{r}
training <- filter(data, Date <= "2017-12-29")
testing <- filter(data, Date > "2017-12-29" & Date <= "2018-12-28")
forecast <-  filter(data, Date > "2018-12-28")
TrainAndTest <- filter(data, Date <= "2018-12-28")
FullPeriod <- filter(data, Date <= "2019-12-27")
```
The data is converted into time series since this portion of the code uses time series econometrics for Maersk B. 
```{r}
#Training
data_train <- xts(training$Genmab, order.by = as.Date(training$Date))
#Test
data_test <- xts(testing$Genmab, order.by = as.Date(testing$Date)) %>% as.data.frame()
#Training+Test
data_traintest <- xts(TrainAndTest$Genmab, order.by = as.Date(TrainAndTest$Date))
#Forecast
data_forecast <- xts(forecast$Genmab, order.by = as.Date(forecast$Date)) %>% as.data.frame()
#Full Period
data_full <- xts(FullPeriod$Genmab, order.by = as.Date(FullPeriod$Date))
```
For convenience the stock is saved as a text for further use. 
```{r}
company <- "Genmab"
```

# Testing for stationarity
To test whether the data set is stationary, we perform ADF-test on the training data. The null hypotesis is the model is statinary.
```{r}
data_train  %>% adf.test() %>% unlist() %>% as.tibble() %>% 
  mutate(Company = company, "Dickey Fuller" = value, "Lag Order" = lead(value), "Alternative Hypothesis" = lead(`Lag Order`), "P-Value" = lead(`Alternative Hypothesis`)) %>% as.data.frame() %>%  head(1) %>% 
  dplyr::select(Company, `Dickey Fuller`, `Lag Order`, `P-Value`, `Alternative Hypothesis`)
```
## First-order differentiating
First-order differentiating the data set and running ADF- test again.
```{r}
data_train  %>% diff() %>% na.omit() %>% adf.test() %>% unlist() %>% as.tibble() %>% mutate(Company = company, "Dickey Fuller" = value, "Lag Order" = lead(value), "Alternative Hypothesis" = lead(`Lag Order`), "P-Value" = lead(`Alternative Hypothesis`)) %>% as.data.frame() %>%  head(1) %>%  dplyr::select(Company, `Dickey Fuller`, `Lag Order`, `P-Value`, `Alternative Hypothesis`)
```
The data set is stationary when it is differentiated one time. The models is a ARIMA(p,1,q). 

# Testing for heteroscedasticity
Using Lagrange Multiplier and Box-Ljung test to test for heteroscedasticity. This is done by using the command *arch.test()*. The command uses Lagrange Multiplier test and a Portmanteau Q test (Box-Ljung test). 

```{r}
data_train  %>% arch.test()  %>% unlist() %>% as.tibble() %>%  mutate(Company = company, "X-squared" = value, df = lead(value), "P-Value" =  lead(df), "Alternative Hypothesis" = lead(lead(lead(`P-Value`))), "P-Value" =ifelse(`P-Value` == 0,  "< 2.2e-16", `P-Value`)) %>% head(1) %>% dplyr::select(Company, `X-squared`, df , `P-Value`, `Alternative Hypothesis`) 
```

# ARIMA

Testing which models have the lowest AIC. This is done by using the command *auto.arima()*.
```{r} 
test_order <- auto.arima(data_train, ic="aic", trace=TRUE) #(0,1,0), (0,1,0) with drift, ARIMA(1,1,0) with drift, ARIMA(0,1,1) with drift.
```
Building the models
```{r}
model1_arima <- Arima(data_train, order=c(0,1,0))
model2_arima <- Arima(data_train, order=c(0,1,0), include.drift=TRUE)
model3_arima <- Arima(data_train, order=c(1,1,0), include.drift=TRUE)
model4_arima <- Arima(data_train, order=c(0,1,1), include.drift=TRUE)

```
out-of-sample forecasting
```{r}
arima_refit1 <- Arima(data_traintest, model=model1_arima) 
arima_fc1 <- window(fitted(arima_refit1), start = 2000) %>% as.data.frame()

arima_refit2 <- Arima(data_traintest, model=model2_arima) 
arima_fc2 <- window(fitted(arima_refit2), start = 2000) %>% as.data.frame()

arima_refit3 <- Arima(data_traintest, model=model3_arima) 
arima_fc3 <- window(fitted(arima_refit3), start = 2000) %>% as.data.frame()

arima_refit4 <- Arima(data_traintest, model=model4_arima) 
arima_fc4 <- window(fitted(arima_refit4), start = 2000) %>% as.data.frame()
```
We combine the forecasted and actual values for simplicity. We rename the variables for clarity.
```{r}
arima_test_fitted1 <- cbind(data_test, arima_fc1) %>% rename(truth = V1, pred = x)
arima_test_fitted2 <- cbind(data_test, arima_fc2) %>% rename(truth = V1, pred = x)
arima_test_fitted3 <- cbind(data_test, arima_fc3) %>% rename(truth = V1, pred = x)
arima_test_fitted4 <- cbind(data_test, arima_fc4) %>% rename(truth = V1, pred = x)
```
Calculating RMSE, MAE, and MAPE.
```{r}
arima_RMSE1 <- arima_test_fitted1 %>% rmse(truth, pred) 
arima_MAE1 <- arima_test_fitted1 %>% mae(truth, pred)
arima_MAPE1 <- arima_test_fitted1 %>% mape(truth, pred) 
arima_eval1 <- cbind(arima_RMSE1$.estimate, arima_MAE1$.estimate, arima_MAPE1$.estimate) %>% as.data.frame() %>% rename("RMSE" = "V1", "MAE" = "V2", "MAPE" = "V3") %>% mutate(Model = "ARIMA(0,1,0)") %>% relocate(Model)
arima_eval1

arima_RMSE2 <- arima_test_fitted2 %>% rmse(truth, pred) 
arima_MAE2 <- arima_test_fitted2 %>% mae(truth, pred)
arima_MAPE2 <- arima_test_fitted2 %>% mape(truth, pred) 
arima_eval2 <- cbind(arima_RMSE2$.estimate, arima_MAE2$.estimate, arima_MAPE2$.estimate) %>% as.data.frame() %>% rename("RMSE" = "V1", "MAE" = "V2", "MAPE" = "V3") %>% mutate(Model = "ARIMA(0,1,0) with drift") %>% relocate(Model)


arima_RMSE3 <- arima_test_fitted3 %>% rmse(truth, pred) 
arima_MAE3 <- arima_test_fitted3 %>% mae(truth, pred)
arima_MAPE3 <- arima_test_fitted3 %>% mape(truth, pred) 
arima_eval3 <- cbind(arima_RMSE3$.estimate, arima_MAE3$.estimate, arima_MAPE3$.estimate) %>% as.data.frame() %>% rename("RMSE" = "V1", "MAE" = "V2", "MAPE" = "V3") %>% mutate(Model = "ARIMA(1,1,0) with drift") %>% relocate(Model)


arima_RMSE4 <- arima_test_fitted4 %>% rmse(truth, pred) 
arima_MAE4 <- arima_test_fitted4 %>% mae(truth, pred)
arima_MAPE4 <- arima_test_fitted4 %>% mape(truth, pred) 
arima_eval4 <- cbind(arima_RMSE4$.estimate, arima_MAE4$.estimate, arima_MAPE4$.estimate) %>% as.data.frame() %>% rename("RMSE" = "V1", "MAE" = "V2", "MAPE" = "V3") %>% mutate(Model = "ARIMA(0,1,1) with drift") %>% relocate(Model)

arima_eval <- rbind(arima_eval1, arima_eval2, arima_eval4, arima_eval3)
arima_eval #Arima(0,1,0)
```
Renameing the models with lowest MAPE, and clearning the enviroment.
```{r}
model_arima <- model1_arima


rm(model1_arima, model2_arima, model3_arima, model4_arima)
```

# ARIMA with features

Creating timeseries for training and train and testing data
The variables are chosen based on Granger test in other R-script
```{r}
variabel1 <- xts(training$Lundbeck, order.by = as.Date(training$Date))
variabel1 <- stats::lag(variabel1) %>% as.data.frame() %>% mutate(`V1` = ifelse(is.na(`V1`), lead(`V1`,1), `V1`)) %>% xts(variabel1$V1, order.by = as.Date(training$Date))

variabel1TT <- xts(TrainAndTest$Lundbeck, order.by = as.Date(TrainAndTest$Date))
variabel1TT <- stats::lag(variabel1TT) %>% as.data.frame() %>% mutate(`V1` = ifelse(is.na(`V1`), lead(`V1`,1), `V1`)) %>% xts(variabel1TT$V1, order.by = as.Date(TrainAndTest$Date))


variabel2 <- xts(training$USD, order.by = as.Date(training$Date))
variabel2 <- stats::lag(variabel2) %>% as.data.frame() %>% mutate(`V1` = ifelse(is.na(`V1`), lead(`V1`,1), `V1`)) %>% xts(variabel2$V1, order.by = as.Date(training$Date)) 

variabel2TT <- xts(TrainAndTest$USD, order.by = as.Date(TrainAndTest$Date))
variabel2TT <- stats::lag(variabel2TT) %>% as.data.frame() %>% mutate(`V1` = ifelse(is.na(`V1`), lead(`V1`,1), `V1`)) %>% xts(variabel2TT$V1, order.by = as.Date(TrainAndTest$Date)) 


variabel3 <- xts(training$SP500, order.by = as.Date(training$Date))
variabel3 <- stats::lag(variabel3) %>% as.data.frame() %>% mutate(`V1` = ifelse(is.na(`V1`), lead(`V1`,1), `V1`)) %>% xts(variabel3$V1, order.by = as.Date(training$Date)) 

variabel3TT <- xts(TrainAndTest$SP500, order.by = as.Date(TrainAndTest$Date))
variabel3TT <- stats::lag(variabel3TT) %>% as.data.frame() %>% mutate(`V1` = ifelse(is.na(`V1`), lead(`V1`,1), `V1`)) %>% xts(variabel3TT$V1, order.by = as.Date(TrainAndTest$Date)) 

```

Bind the variables together

```{r}
variables_train <- cbind(variabel1, variabel2, variabel3)
variables_TT <- cbind(variabel1TT, variabel2TT, variabel3TT)

rm(variabel1, variabel2, variabel3, variabel1TT, variabel2TT, variabel3TT)
```

Building the model

```{r}
model1_arima <- auto.arima(data_train, ic="aic", xreg = variables_train, trace=T)
summary(model1_arima)
```

```{r}
model1_arima_features <- Arima(data_train, order=c(0,1,0), xreg = variables_train)
model2_arima_features <- Arima(data_train, order=c(2,1,2), xreg = variables_train)
model3_arima_features <- Arima(data_train, order=c(0,1,1), xreg = variables_train)
model4_arima_features <- Arima(data_train, order=c(1,1,0), xreg = variables_train)
```


Creating out of sample forecasts

```{r}
arima_refit1 <- Arima(data_traintest, model=model1_arima_features, xreg = variables_TT)
arima_fc1 <- window(fitted(arima_refit1), start = 2000) %>% as.data.frame()

arima_refit2 <- Arima(data_traintest, model=model2_arima_features, xreg = variables_TT)
arima_fc2 <- window(fitted(arima_refit2), start = 2000) %>% as.data.frame()


arima_refit3 <- Arima(data_traintest, model=model3_arima_features, xreg = variables_TT)
arima_fc3 <- window(fitted(arima_refit3), start = 2000) %>% as.data.frame()

arima_refit4 <- Arima(data_traintest, model=model4_arima_features, xreg = variables_TT)
arima_fc4 <- window(fitted(arima_refit4), start = 2000) %>% as.data.frame()
```

Bind actual and predicted values together

```{r}
arima_test_fitted1 <- cbind(data_test, arima_fc1) %>% rename(truth = V1, pred = x)
arima_test_fitted2 <- cbind(data_test, arima_fc2) %>% rename(truth = V1, pred = x)
arima_test_fitted3 <- cbind(data_test, arima_fc3) %>% rename(truth = V1, pred = x)
arima_test_fitted4 <- cbind(data_test, arima_fc4) %>% rename(truth = V1, pred = x)
```


Calcualateing Evaluation stats

```{r}
arima_RMSE1 <- arima_test_fitted1 %>% rmse(truth, pred) 
arima_MAE1 <- arima_test_fitted1 %>% mae(truth, pred)
arima_MAPE1 <- arima_test_fitted1 %>% mape(truth, pred) 
arima_eval1 <- cbind(arima_RMSE1$.estimate, arima_MAE1$.estimate, arima_MAPE1$.estimate) %>% as.data.frame() %>% rename("RMSE" = "V1", "MAE" = "V2", "MAPE" = "V3") %>% mutate(Model = "ARIMA(0,1,0) with Features") %>% relocate(Model)
arima_eval1

arima_RMSE2 <- arima_test_fitted2 %>% rmse(truth, pred) 
arima_MAE2 <- arima_test_fitted2 %>% mae(truth, pred)
arima_MAPE2 <- arima_test_fitted2 %>% mape(truth, pred) 
arima_eval2 <- cbind(arima_RMSE2$.estimate, arima_MAE2$.estimate, arima_MAPE2$.estimate) %>% as.data.frame() %>% rename("RMSE" = "V1", "MAE" = "V2", "MAPE" = "V3") %>% mutate(Model = "ARIMA(2,1,2) with Features") %>% relocate(Model)


arima_RMSE3 <- arima_test_fitted3 %>% rmse(truth, pred) 
arima_MAE3 <- arima_test_fitted3 %>% mae(truth, pred)
arima_MAPE3 <- arima_test_fitted3 %>% mape(truth, pred) 
arima_eval3 <- cbind(arima_RMSE3$.estimate, arima_MAE3$.estimate, arima_MAPE3$.estimate) %>% as.data.frame() %>% rename("RMSE" = "V1", "MAE" = "V2", "MAPE" = "V3") %>% mutate(Model = "ARIMA(0,1,1) with Features") %>% relocate(Model)


arima_RMSE4 <- arima_test_fitted4 %>% rmse(truth, pred) 
arima_MAE4 <- arima_test_fitted4 %>% mae(truth, pred)
arima_MAPE4 <- arima_test_fitted4 %>% mape(truth, pred) 
arima_eval4 <- cbind(arima_RMSE4$.estimate, arima_MAE4$.estimate, arima_MAPE4$.estimate) %>% as.data.frame() %>% rename("RMSE" = "V1", "MAE" = "V2", "MAPE" = "V3") %>% mutate(Model = "ARIMA(1,1,0) with Features") %>% relocate(Model)


arima_eval <- rbind(arima_eval, arima_eval2, arima_eval1, arima_eval3, arima_eval4 ) %>% arrange(MAPE)
arima_eval  #Arima(1,1,0) with Features

```

# Forecast

First we choose the model from the out of sample forecasts with the lowest MAPE, in this case is it the ARIMAX(1,1,0) (model4_arima_features)


```{r}
model_Genmab <- model4_arima_features
```


Creating xreg for the full period.
```{r}
variabel1FC <- xts(FullPeriod$Lundbeck, order.by = as.Date(FullPeriod$Date))
variabel1FC <- stats::lag(variabel1FC) %>% as.data.frame() %>% mutate(`V1` = ifelse(is.na(`V1`), lead(`V1`,1), `V1`)) %>% xts(variabel1FC$V1, order.by = as.Date(FullPeriod$Date))

variabel2FC <- xts(FullPeriod$USD, order.by = as.Date(FullPeriod$Date))
variabel2FC <- stats::lag(variabel2FC) %>% as.data.frame() %>% mutate(`V1` = ifelse(is.na(`V1`), lead(`V1`,1), `V1`)) %>% xts(variabel2FC$V1, order.by = as.Date(FullPeriod$Date))

variabel3FC <- xts(FullPeriod$SP500, order.by = as.Date(FullPeriod$Date))
variabel3FC <- stats::lag(variabel3FC) %>% as.data.frame() %>% mutate(`V1` = ifelse(is.na(`V1`), lead(`V1`,1), `V1`)) %>% xts(variabel3FC$V1, order.by = as.Date(FullPeriod$Date))


variables_FC <- cbind(variabel1FC, variabel2FC, variabel3FC)

rm(variabel1FC, variabel2FC, variabel3FC)
```

Creating the forecasts using the best model

```{r}
arima_refit_forecast <- Arima(data_full, model=model_Genmab, xreg = variables_FC)
arima_fc_forecast <- window(fitted(arima_refit_forecast), start = 2248) %>% as.data.frame()
```

Binding predicted and actual values together
```{r}
arima_actual_forecasts <- cbind(data_forecast, arima_fc_forecast) %>% rename(truth = V1, pred = x)
```

```{r}
arima_actual_forecasts %>% write.csv("ARIMA_Genmab_Forecast_Series")
```


Calculateing evaluation statistics

```{r}
arima_RMSE_forecast <- arima_actual_forecasts %>% rmse(truth, pred) 
arima_MAE_forecast <- arima_actual_forecasts %>% mae(truth, pred)
arima_MAPE_forecast <- arima_actual_forecasts %>% mape(truth, pred) 
arima_eval_forecast <- cbind(arima_RMSE_forecast$.estimate, arima_MAE_forecast$.estimate, arima_MAPE_forecast$.estimate) %>% as.data.frame() %>% rename("RMSE" = "V1", "MAE" = "V2", "MAPE" = "V3") #%>% mutate(Model = "ARIMA(0,0,28) with Features") %>% relocate(Model)
arima_eval_forecast
```