---
title: "Econometrics - SomCorp"
author: "Andreas Borup Joergensen, Mette Koch Moeller, Robert Hoestrup"
date: "04-03-2021"
output:
  html_notebook:
    code_folding: show
    df_print: paged
    theme: flatly
    toc: yes
    toc_depth: 4
    toc_float:
      collapsed: no
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '2'
---
```{r setup, include=FALSE}
### Generic preamble
rm(list=ls())
Sys.setenv(LANG = "en")
options(scipen = 5)

library(knitr)
knitr::opts_chunk$set(warning=FALSE,
                     message=FALSE,
                     comment=FALSE, 
                     fig.align="center"
                     )
```

Downloading the packages
```{r}
library(tidyverse) #Collection of all the good stuff like dplyr, ggplot2 ect.
library(magrittr) #For advanced piping
library(ggpubr) #package that helps mixing multiple plots on the same page
library(wooldridge)
library(car)
library(lmtest)
library(sandwich)
library(quantmod)
library(tseries)
library(urca)
library(vars)
library(forecast)
library(nortsTest)
library(yardstick)
library(midasr) # Brugt til static forecasting
#'forecast', 'midasr', 'nortsTest', 'quantmod', 'sandwich', 'tseries', 'urca', 'vars', 'wooldridge'
```

Loading data from GitHub

```{r} 
data <- read.csv("https://raw.githubusercontent.com/andreasbj77/Master-thesis/main/Data/Final_data/DataCreationSimCorp.csv")
```

# Preprocessing
We split up our data set into training, testing, training+testing and forecasting. Splitting up the data makes it possible for us to make out-of-sample forecasting. The training+testing data frame is created so we can make one step ahead forecasts in the testing period.
```{r}
training <- filter(data, Date <= "2017-12-29")
testing <- filter(data, Date > "2017-12-29" & Date <= "2018-12-28")
forecast <-  filter(data, Date > "2018-12-28")
TrainAndTest <- filter(data, Date <= "2018-12-28")
FullPeriod <- filter(data, Date <= "2019-12-27")
```
The data is converted into time series since this portion of the code uses time series econometrics for Maersk B. 
```{r}
#Training
data_train <- xts(training$SimCorp, order.by = as.Date(training$Date))
#Test
data_test <- xts(testing$SimCorp, order.by = as.Date(testing$Date)) %>% as.data.frame()
#Training+Test
data_traintest <- xts(TrainAndTest$SimCorp, order.by = as.Date(TrainAndTest$Date))
#Forecast
data_forecast <- xts(forecast$SimCorp, order.by = as.Date(forecast$Date)) %>% as.data.frame()
#Full Period
data_full <- xts(FullPeriod$SimCorp, order.by = as.Date(FullPeriod$Date))
```
For convenience the stock is saved as a text for further use. 
```{r}
company <- "SimCorp"
```

# Testing for stationarity
To test whether the data set is stationary, we perform ADF-test on the training data. The null hypotesis is the model is statinary.
```{r}
data_train  %>% adf.test() %>% unlist() %>% as.tibble() %>% 
  mutate(Company = company, "Dickey Fuller" = value, "Lag Order" = lead(value), "Alternative Hypothesis" = lead(`Lag Order`), "P-Value" = lead(`Alternative Hypothesis`)) %>% as.data.frame() %>%  head(1) %>% 
  dplyr::select(Company, `Dickey Fuller`, `Lag Order`, `P-Value`, `Alternative Hypothesis`)
```
## First-order differentiating
First-order differentiating the data set and running ADF- test again.
```{r}
data_train  %>% diff() %>% na.omit() %>% adf.test() %>% unlist() %>% as.tibble() %>% mutate(Company = company, "Dickey Fuller" = value, "Lag Order" = lead(value), "Alternative Hypothesis" = lead(`Lag Order`), "P-Value" = lead(`Alternative Hypothesis`)) %>% as.data.frame() %>%  head(1) %>%  dplyr::select(Company, `Dickey Fuller`, `Lag Order`, `P-Value`, `Alternative Hypothesis`)
```
The data set is stationary when it is differentiated one time. The models is a ARIMA(p,1,q). 

# Testing for heteroscedasticity
Using Lagrange Multiplier and Box-Ljung test to test for heteroscedasticity. This is done by using the command *arch.test()*. The command uses Lagrange Multiplier test and a Portmanteau Q test (Box-Ljung test).  

```{r}
data_train  %>% arch.test()  %>% unlist() %>% as.tibble() %>%  mutate(Company = company, "X-squared" = value, df = lead(value), "P-Value" =  lead(df), "Alternative Hypothesis" = lead(lead(lead(`P-Value`))), "P-Value" =ifelse(`P-Value` == 0,  "< 2.2e-16", `P-Value`)) %>% head(1) %>% dplyr::select(Company, `X-squared`, df , `P-Value`, `Alternative Hypothesis`) 
```
# ARIMA

Testing which models have the lowest AIC. This is done by using the command *auto.arima()*.
```{r} 
test_order <- auto.arima(data_train,  ic="aic", trace=TRUE) #(3,1,0), (3,1,0) with drift, ARIMA(3,1,1), ARIMA(2,1,1).

```

```{r}
model1_arima <- Arima(data_train, order=c(0,1,0))
model2_arima <- Arima(data_train, order=c(0,1,0), include.drift=TRUE)
model3_arima <- Arima(data_train, order=c(0,1,1), include.drift=TRUE)
model4_arima <- Arima(data_train, order=c(1,1,0), include.drift=TRUE)

```
out-of-sample forecasting
```{r}
arima_refit1 <- Arima(data_traintest, model=model1_arima) 
arima_fc1 <- window(fitted(arima_refit1), start = 2000) %>% as.data.frame()

arima_refit2 <- Arima(data_traintest, model=model2_arima) 
arima_fc2 <- window(fitted(arima_refit2), start = 2000) %>% as.data.frame()

arima_refit3 <- Arima(data_traintest, model=model3_arima) 
arima_fc3 <- window(fitted(arima_refit3), start = 2000) %>% as.data.frame()

arima_refit4 <- Arima(data_traintest, model=model4_arima) 
arima_fc4 <- window(fitted(arima_refit4), start = 2000) %>% as.data.frame()
```
We combine the forecasted and actual values for simplicity. We rename the variables for clarity.
```{r}
arima_test_fitted1 <- cbind(data_test, arima_fc1) %>% rename(truth = V1, pred = x)
arima_test_fitted2 <- cbind(data_test, arima_fc2) %>% rename(truth = V1, pred = x)
arima_test_fitted3 <- cbind(data_test, arima_fc3) %>% rename(truth = V1, pred = x)
arima_test_fitted4 <- cbind(data_test, arima_fc4) %>% rename(truth = V1, pred = x)
```
Calculating RMSE, MAE, and MAPE.
```{r}
arima_RMSE1 <- arima_test_fitted1 %>% rmse(truth, pred) 
arima_MAE1 <- arima_test_fitted1 %>% mae(truth, pred)
arima_MAPE1 <- arima_test_fitted1 %>% mape(truth, pred) 
arima_eval1 <- cbind(arima_RMSE1$.estimate, arima_MAE1$.estimate, arima_MAPE1$.estimate) %>% as.data.frame() %>% rename("RMSE" = "V1", "MAE" = "V2", "MAPE" = "V3") %>% mutate(Model = "ARIMA(0,1,0)") %>% relocate(Model)
arima_eval1

arima_RMSE2 <- arima_test_fitted2 %>% rmse(truth, pred) 
arima_MAE2 <- arima_test_fitted2 %>% mae(truth, pred)
arima_MAPE2 <- arima_test_fitted2 %>% mape(truth, pred) 
arima_eval2 <- cbind(arima_RMSE2$.estimate, arima_MAE2$.estimate, arima_MAPE2$.estimate) %>% as.data.frame() %>% rename("RMSE" = "V1", "MAE" = "V2", "MAPE" = "V3") %>% mutate(Model = "ARIMA(0,1,0) with drift") %>% relocate(Model)


arima_RMSE3 <- arima_test_fitted3 %>% rmse(truth, pred) 
arima_MAE3 <- arima_test_fitted3 %>% mae(truth, pred)
arima_MAPE3 <- arima_test_fitted3 %>% mape(truth, pred) 
arima_eval3 <- cbind(arima_RMSE3$.estimate, arima_MAE3$.estimate, arima_MAPE3$.estimate) %>% as.data.frame() %>% rename("RMSE" = "V1", "MAE" = "V2", "MAPE" = "V3") %>% mutate(Model = "ARIMA(0,1,1) with drift") %>% relocate(Model)


arima_RMSE4 <- arima_test_fitted4 %>% rmse(truth, pred) 
arima_MAE4 <- arima_test_fitted4 %>% mae(truth, pred)
arima_MAPE4 <- arima_test_fitted4 %>% mape(truth, pred) 
arima_eval4 <- cbind(arima_RMSE4$.estimate, arima_MAE4$.estimate, arima_MAPE4$.estimate) %>% as.data.frame() %>% rename("RMSE" = "V1", "MAE" = "V2", "MAPE" = "V3") %>% mutate(Model = "ARIMA(1,1,0) with drift") %>% relocate(Model)

arima_eval <- rbind(arima_eval2, arima_eval1, arima_eval3, arima_eval4) %>% arrange(MAPE)
arima_eval
```

# ARIMA with features

Creating timeseries for training and train and testing data.


Creating timeseries for training and train and testing data.The variables are chosen based on Granger test in other R-script.
We lag the series which turns the first observation into NA. we replace the NA with the second observation.

```{r}

variabel1 <- xts(training$DSV, order.by = as.Date(training$Date))
variabel1 <- stats::lag(variabel1) %>% as.data.frame() %>% mutate(`V1` = ifelse(is.na(`V1`), lead(`V1`,1), `V1`)) %>% xts(variabel1$V1, order.by = as.Date(training$Date))

variabel1TT <- xts(TrainAndTest$DSV, order.by = as.Date(TrainAndTest$Date))
variabel1TT <- stats::lag(variabel1TT) %>% as.data.frame() %>% mutate(`V1` = ifelse(is.na(`V1`), lead(`V1`,1), `V1`)) %>% xts(variabel1TT$V1, order.by = as.Date(TrainAndTest$Date)) 


variabel2 <- xts(training$Genmab, order.by = as.Date(training$Date))
variabel2 <- stats::lag(variabel2) %>% as.data.frame() %>% mutate(`V1` = ifelse(is.na(`V1`), lead(`V1`,1), `V1`)) %>% xts(variabel2$V1, order.by = as.Date(training$Date))

variabel2TT <- xts(TrainAndTest$Genmab, order.by = as.Date(TrainAndTest$Date))
variabel2TT <- stats::lag(variabel2TT) %>% as.data.frame() %>% mutate(`V1` = ifelse(is.na(`V1`), lead(`V1`,1), `V1`)) %>% xts(variabel2TT$V1, order.by = as.Date(TrainAndTest$Date)) 


variabel3 <- xts(training$SydBank, order.by = as.Date(training$Date))
variabel3 <- stats::lag(variabel3) %>% as.data.frame() %>% mutate(`V1` = ifelse(is.na(`V1`), lead(`V1`,1), `V1`)) %>% xts(variabel3$V1, order.by = as.Date(training$Date)) 

variabel3TT <- xts(TrainAndTest$SydBank, order.by = as.Date(TrainAndTest$Date))
variabel3TT <- stats::lag(variabel3TT) %>% as.data.frame() %>% mutate(`V1` = ifelse(is.na(`V1`), lead(`V1`,1), `V1`)) %>% xts(variabel3TT$V1, order.by = as.Date(TrainAndTest$Date)) 


variabel4 <- xts(training$WilliamDemant, order.by = as.Date(training$Date))
variabel4 <- stats::lag(variabel4) %>% as.data.frame() %>% mutate(`V1` = ifelse(is.na(`V1`), lead(`V1`,1), `V1`)) %>% xts(variabel4$V1, order.by = as.Date(training$Date))

variabel4TT <- xts(TrainAndTest$WilliamDemant, order.by = as.Date(TrainAndTest$Date))
variabel4TT <- stats::lag(variabel4TT) %>% as.data.frame() %>% mutate(`V1` = ifelse(is.na(`V1`), lead(`V1`,1), `V1`)) %>% xts(variabel4TT$V1, order.by = as.Date(TrainAndTest$Date)) 


variabel5 <- xts(training$SP500, order.by = as.Date(training$Date))
variabel5 <- stats::lag(variabel5) %>% as.data.frame() %>% mutate(`V1` = ifelse(is.na(`V1`), lead(`V1`,1), `V1`)) %>% xts(variabel5$V1, order.by = as.Date(training$Date)) 

variabel5TT <- xts(TrainAndTest$SP500, order.by = as.Date(TrainAndTest$Date))
variabel5TT <- stats::lag(variabel5TT) %>% as.data.frame() %>% mutate(`V1` = ifelse(is.na(`V1`), lead(`V1`,1), `V1`)) %>% xts(variabel5TT$V1, order.by = as.Date(TrainAndTest$Date))

```

Bind the variables together

```{r}
variables_train <- cbind(variabel1, variabel2, variabel3, variabel4, variabel5)
variables_TT <- cbind(variabel1TT, variabel2TT, variabel3TT, variabel4TT, variabel5TT)


rm(variabel1, variabel2, variabel3, variabel4, variabel5, variabel1TT, variabel2TT, variabel3TT, variabel4TT, variabel5TT)

```

Building the model

```{r}
model1_arima <- auto.arima(data_train, xreg = variables_train, trace=T)
summary(model1_arima)
```

```{r}
model1_arima <- Arima(data_train, order=c(1,1,1), xreg = variables_train)
model2_arima <- Arima(data_train, order=c(0,1,0), xreg = variables_train)
model3_arima <- Arima(data_train, order=c(0,1,1), xreg = variables_train)
model4_arima <- Arima(data_train, order=c(1,1,0), xreg = variables_train)
```



Creating out of sample forecasts

```{r}
arima_refit1 <- Arima(data_traintest, model=model1_arima, xreg = variables_TT)
arima_fc1 <- window(fitted(arima_refit1), start = 2000) %>% as.data.frame()

arima_refit2 <- Arima(data_traintest, model=model2_arima, xreg = variables_TT)
arima_fc2 <- window(fitted(arima_refit2), start = 2000) %>% as.data.frame()

arima_refit3 <- Arima(data_traintest, model=model3_arima, xreg = variables_TT)
arima_fc3 <- window(fitted(arima_refit3), start = 2000) %>% as.data.frame()

arima_refit4 <- Arima(data_traintest, model=model4_arima, xreg = variables_TT)
arima_fc4 <- window(fitted(arima_refit4), start = 2000) %>% as.data.frame()
```

Bind actual and predicted values together

```{r}
arima_test_fitted1 <- cbind(data_test, arima_fc1) %>% rename(truth = V1, pred = x)
arima_test_fitted2 <- cbind(data_test, arima_fc2) %>% rename(truth = V1, pred = x)
arima_test_fitted3 <- cbind(data_test, arima_fc3) %>% rename(truth = V1, pred = x)
arima_test_fitted4 <- cbind(data_test, arima_fc4) %>% rename(truth = V1, pred = x)
```


Calcualateing Evaluation stats

```{r}
arima_RMSE1 <- arima_test_fitted1 %>% rmse(truth, pred) 
arima_MAE1 <- arima_test_fitted1 %>% mae(truth, pred)
arima_MAPE1 <- arima_test_fitted1 %>% mape(truth, pred) 
arima_eval1 <- cbind(arima_RMSE1$.estimate, arima_MAE1$.estimate, arima_MAPE1$.estimate) %>% as.data.frame() %>% rename("RMSE" = "V1", "MAE" = "V2", "MAPE" = "V3") %>% mutate(Model = "ARIMA(1,1,1) with Features") %>% relocate(Model)
arima_eval1

arima_RMSE2 <- arima_test_fitted2 %>% rmse(truth, pred) 
arima_MAE2 <- arima_test_fitted2 %>% mae(truth, pred)
arima_MAPE2 <- arima_test_fitted2 %>% mape(truth, pred) 
arima_eval2 <- cbind(arima_RMSE2$.estimate, arima_MAE2$.estimate, arima_MAPE2$.estimate) %>% as.data.frame() %>% rename("RMSE" = "V1", "MAE" = "V2", "MAPE" = "V3") %>% mutate(Model = "ARIMA(0,1,0) with Features") %>% relocate(Model)


arima_RMSE3 <- arima_test_fitted3 %>% rmse(truth, pred) 
arima_MAE3 <- arima_test_fitted3 %>% mae(truth, pred)
arima_MAPE3 <- arima_test_fitted3 %>% mape(truth, pred) 
arima_eval3 <- cbind(arima_RMSE3$.estimate, arima_MAE3$.estimate, arima_MAPE3$.estimate) %>% as.data.frame() %>% rename("RMSE" = "V1", "MAE" = "V2", "MAPE" = "V3") %>% mutate(Model = "ARIMA(0,1,1) with Features") %>% relocate(Model)


arima_RMSE4 <- arima_test_fitted4 %>% rmse(truth, pred) 
arima_MAE4 <- arima_test_fitted4 %>% mae(truth, pred)
arima_MAPE4 <- arima_test_fitted4 %>% mape(truth, pred) 
arima_eval4 <- cbind(arima_RMSE4$.estimate, arima_MAE4$.estimate, arima_MAPE4$.estimate) %>% as.data.frame() %>% rename("RMSE" = "V1", "MAE" = "V2", "MAPE" = "V3") %>% mutate(Model = "ARIMA(1,1,0) with Features") %>% relocate(Model)

arima_eval <- rbind(arima_eval, arima_eval1, arima_eval2, arima_eval3, arima_eval4) %>% arrange(MAPE)
arima_eval  #Arima(0,1,0)
```

# Forecast

First we choose the model from the out of sample forecasts with the lowest MAPE and RMSE, in this case point RMSE at one model and MAPE at another. In this case we, then choose the model of the two with the lowest AIC which is the ARIMAX(0,1,0) (model2_arima)


```{r}
model_SimCorp<- model2_arima
```

Creating xreg for the full period.
```{r}
variabel1FC <- xts(FullPeriod$DSV, order.by = as.Date(FullPeriod$Date))
variabel1FC <- stats::lag(variabel1FC) %>% as.data.frame() %>% mutate(`V1` = ifelse(is.na(`V1`), lead(`V1`,1), `V1`)) %>% xts(variabel1FC$V1, order.by = as.Date(FullPeriod$Date))

variabel2FC <- xts(FullPeriod$Genmab, order.by = as.Date(FullPeriod$Date))
variabel2FC <- stats::lag(variabel2FC) %>% as.data.frame() %>% mutate(`V1` = ifelse(is.na(`V1`), lead(`V1`,1), `V1`)) %>% xts(variabel2FC$V1, order.by = as.Date(FullPeriod$Date))

variabel3FC <- xts(FullPeriod$SydBank, order.by = as.Date(FullPeriod$Date))
variabel3FC <- stats::lag(variabel3FC) %>% as.data.frame() %>% mutate(`V1` = ifelse(is.na(`V1`), lead(`V1`,1), `V1`)) %>% xts(variabel3FC$V1, order.by = as.Date(FullPeriod$Date))

variabel4FC <- xts(FullPeriod$WilliamDemant, order.by = as.Date(FullPeriod$Date))
variabel4FC <- stats::lag(variabel4FC) %>% as.data.frame() %>% mutate(`V1` = ifelse(is.na(`V1`), lead(`V1`,1), `V1`)) %>% xts(variabel4FC$V1, order.by = as.Date(FullPeriod$Date))

variabel5FC <- xts(FullPeriod$SP500, order.by = as.Date(FullPeriod$Date))
variabel5FC <- stats::lag(variabel5FC) %>% as.data.frame() %>% mutate(`V1` = ifelse(is.na(`V1`), lead(`V1`,1), `V1`)) %>% xts(variabel5FC$V1, order.by = as.Date(FullPeriod$Date))

variables_FC <- cbind(variabel1FC, variabel2FC, variabel3FC, variabel4FC, variabel5FC)

rm(variabel1FC, variabel2FC, variabel3FC, variabel4FC, variabel5FC)
```

Creating the forecasts using the best model

```{r}
arima_refit_forecast <- Arima(data_full, model=model_SimCorp, xreg = variables_FC)
arima_fc_forecast <- window(fitted(arima_refit_forecast), start = 2248) %>% as.data.frame()
```

Binding predicted and actual values together
```{r}
arima_actual_forecasts <- cbind(data_forecast, arima_fc_forecast) %>% rename(truth = V1, pred = x)
```


```{r}
arima_actual_forecasts %>% write.csv("ARIMA_SimCorp_Forecast_Series")
```


Calculateing evaluation statistics

```{r}
arima_RMSE_forecast <- arima_actual_forecasts %>% rmse(truth, pred) 
arima_MAE_forecast <- arima_actual_forecasts %>% mae(truth, pred)
arima_MAPE_forecast <- arima_actual_forecasts %>% mape(truth, pred) 
arima_eval_forecast <- cbind(arima_RMSE_forecast$.estimate, arima_MAE_forecast$.estimate, arima_MAPE_forecast$.estimate) %>% as.data.frame() %>% rename("RMSE" = "V1", "MAE" = "V2", "MAPE" = "V3") #%>% mutate(Model = "ARIMA(0,0,28) with Features") %>% relocate(Model)
arima_eval_forecast
```